\textbf{Computational Fuzzy Extractors}

### 2013-05-20 CompFuzzExt.tex
typo fixed

\title{\textbf{Computational Fuzzy Extractors}}
\author{Benjamin Fuller%\footnote{The work of Benjamin Fuller is sponsored by the United States Air Force under Air Force Contract FA8721-05-C-0002. Opinions, interpretations, conclusions and recommendations are those of the authors and are not necessarily endorsed by the United States Government.}
 \and Xianrui Meng \and Leonid Reyzin}
\begin{document}
\maketitle

@@

\item\textbf{Positive Result:} We show that the negative result can be overcome by
analyzing computational fuzzy extractors directly.  Namely, we show
how to build a computational fuzzy extractor whose output key length
equals the entropy of the source (this is impossible in the
information-theoretic setting). Our construction is based on the
hardness of the Learning with Errors (LWE) problem, and is secure when
the noisy source is uniform or symbol-fixing (that is, each dimension
is either uniform or fixed). As part of the security proof, we show a result of independent interest, namely
that the decision version of LWE is secure even when a small number of
dimensions has no error.
\end{itemize}
\textbf{Keywords} Fuzzy extractors, secure sketches, key derivation, Learning with Errors, noise tolerance, error-correcting codes, computational entropy, randomness extractors.
\end{abstract}

@@

The rest of this section is devoted to making these intuitive statements precise.
 We describe the \class{LWE} problem and the security of our construction in \secref{subsec:LWE}.
 %a recent variant due to D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} in \secref{subsec:LWE}.  
We describe one possible polynomial-time $\decode_t$ (which corrects more errors than is possible by exhaustive search) in \secref{sec:time main construction}.  In \secref{sec:lossless extractor}, we describe parameter settings where we extract as many bits as the size of our input distribution and thus our construction is lossless.  In \secref{sec:prg based comparison}, we compare \consref{cons:informal construction} to using a sketch-and-extract approach (\lemref{lem:fuzzy ext construction}) instantiated with a computational extractor.

\subsection{Security of \consref{cons:informal construction}}
\label{subsec:LWE}
The $\LWE$ problem was introduced by Regev \cite{regev2005LWE, regevLWEsurvey} as a generalization of ``learning parity with noise." For a positive integers $n$,  a prime $q$, a vector $\vx \in \Fq^n$, and a probability distribution $\chi$ on $\Fq$, let $A_{\vect{x}, \chi}$ be the distribution obtained by choosing $\vA \overset{\$}\leftarrow \Fq^{m\times n}$ uniformly at random and a noise term $\ve \overset{\$}\leftarrow \chi^m$, and outputting $(\vA, \vA\vx+\ve)$. 
%The Learning With Errors ($\lwe$) is defined as follows: given access to the oracle that outputs a sample from the distribution $A_{\vx, \chi}$ for a uniform $\vx$, and output $\vx$ with high probability.

\begin{definition}[Decisional $\lwe$]\label{def:dist-LWE}Let $n$ be a security parameter.  
Let $m = m(n) = \poly(n)$ be an integer $q = q(n) = \poly(n)$ be a prime. The decisional version of the $\LWE$ problem, denoted \class{dist}-$\LWE_{n, m, q, \chi}$, is to distinguish between $m$ samples chosen according to $A_{\vx, \chi}$ for a uniformly random $\vx$ from samples chosen according to the uniform distribution over $(\Fq^{m\times n}, \Fq^m)$.

We say that $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})$-secure if no (probabilistic) distinguisher of size $s_{sec}$ can distinguish the $\lwe$ instances from uniform except with probability $\epsilon$.  If $\distLWE_{n, m, q, \chi}$ is secure for all $s_{sec} = \poly(n), \epsilon  = 1/\poly(n)$, we say it is \emph{secure}.
\end{definition}

Denote by $\bar{\Psi}_\alpha$ the discretized Gaussian distribution over $\Fq$ with variance $(\alpha q)^2/(2\pi)$. Regev\cite{regev2005LWE} and Peikert \cite{peikert2009latticereduction} show that, when $\chi = \bar{\Psi}_{\alpha}$, $\class{dist}$-$\lwe_{n, m, q, \chi}$ is secure if the worst-case lattice problems are hard to approximate using quantum algorithms.
%The matrix $\vA$ and secret $\vx$ are drawn from the uniform distribution, while the error term is drawn from a discretized Gaussian distribution $\bar{\Psi}_{\alpha}$.  
Furthermore, Regev \cite{regev2005LWE} shows a decision to search equivalence for $q=\poly(n)$.
%We state the decisional version in terms of unpredictability entropy~(see \defref{def:unp entropy} for the definition of unpredictability entropy and \secref{sec:LWEoverview} for an equivalence to the standard formulation of decisional LWE):

\begin{proposition} 
\label{assume:entropy LWE}
Let $n$ be a security parameter, let $m = m(n) = \poly(n)$ be an integer and $q = q(n) = \poly(n)$ be a prime and let $\bar{\Psi}_\alpha$ be the discretized Gaussian distribution with variance $(\alpha q)^2/2\pi$.  If GAPSVP and SIVP are hard to approximate~(on lattices of dimension $n$) within polynomial factors for quantum algorithms, then $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})$-secure for $\epsilon = \ngl(n)$ and $s_{sec} = \poly(n)$.  That is, $\delta^{\mathcal{D}_{s_{sec}}}(\vA, \vA X+E ), (\vA, U_{m\log q}))\leq \epsilon$.

@@

\subsection{Efficiency of \consref{cons:informal construction}}
\label{sec:time main construction}
%\subsection{Analysis of Construction~\ref{cons:LWESecureSketch}}
\consref{cons:informal construction} is only useful if $\decode_t$ can be efficiently implemented.  We need a decoding algorithm for a random linear code with $t$ errors that runs in polynomial time.  We present a simple $\decode_t$.  Replacing our algorithm  with a state of the art algorithm will only increase our correcting capability and improve \consref{cons:informal construction}.

\begin{construction}
\label{cons:decoding algorithm} We consider a setting of $(n, m, q, \chi)$ where $m\geq 3n$.  We describe $\decode_t$:
\begin{enumerate}
\item Input $\vA , B = \vA X + W - W'$
\item Randomly select rows without replacement $i_1,..., i_{2n}\leftarrow [1,m]$.  
\item Restrict $\vA, B$ to rows $i_1,...,i_{2n}$, denote this as $\vA_{i_1,...,i_{2n}}, B_{i_1,...,i_{2n}}$.
\item Find $n$ rows of $\vA_{i_1,..., i_{2n}}$ that are linearly independent.  
If no such rows exist go to step (2).
\item Denote by $\vA', B'$ the restriction of $\vA_{i_1,..., i_{2n}}, B_{i_1,..., i_{2n}}$ (respectively) to these rows. Compute $X' = (\vA')^{-1}B'$.  If $B- AX'$ has more than $t$ nonzero coordinates, go to step (2).
\item Output $X'$.
\end{enumerate}
\end{construction}

$\decode_t$ is an efficient decoding procedure when $t$ is small enough.  We need $t$ to be small enough for none of the selected rows to have errors with probability at least $\frac{1}{\poly(n)}$. Furthermore, we need the code generated by $\vA$ to have distance at least $t$ with high probability.  The second condition is to ensure that there are few false positives in step (5).  This condition is satisfied for $t = O(\frac{m}{n}\log n)$, since random codes have high distance with probability $1-e^{-\Omega(n)}$.  The exact statement is in \corref{cor:code high distance}.  Furthermore, steps (3-5) are computable in time $O(n^3)$.
This is also a sufficient condition for steps (3-5) to succeed with probability at least $1/\poly(n)$.  We consider the setting where the running time of $\decode_t$ grows with $t$.  $\decode_t$ is efficient as long as $t$ is small enough.  The formal statement is below~(proof in \secref{sec:proof lem i t poly time}):
%  We provide the results in \lemref{lem:i t constant time} and \lemref{lem:i t poly time} respectively~(proofs in Sections~\ref{sec:proof lem i t constant time},~\ref{sec:proof lem i t poly time}).  \lnote{can we combine these lemmas into one?  It's hard for the reader to compare line-by-line to see where the differences are.  More errors seems better; we could simply point out that constant number of erros leads to better running time, right in the statement of the lemma or right after it}
%\begin{lemma}[Efficiency of $\decode_t$ when $t\leq m/n-1$]
%\label{lem:i t constant time}
%Let $\vA, \vA X+E$ be a  $(n,m, q, \chi)$-LWE instance that is a unique witness relation~(except with negligible probability).  Assume $d(E, E')\leq t$ where $t \leq (m-n)/n$.  Then $\decode_t$ runs in expected time $O(n^3\log q+ s_{ver})$ on input distribution $\vA, \vA X+E - E'$ and outputs $X' =X$ for all but a negligible fraction of inputs over $\vA, \vA X+E - E'$.
%\end{lemma}
\begin{lemma}[Efficiency of $\decode_t$ when $t\leq d\log n (m/n-2)$]
\label{lem:i t poly time}
Let $\vA, \vA X+W$ be a  $\distLWE_{n,m, q, \chi}$ instance where $m\geq 3n$.  Assume that $W$ is split into $m$ blocks of length $b$ and let $\dis$ be the Hamming distance over alphabet $2^b$.  Let $d$ be a constant and assume that $\dis(W, W')\leq t$ where $t\leq d\log n(\frac{m}{n}-2)$.  Then $\decode_t$ runs in expected time $O(n^{4d}(n^3\log q))$ on input distribution $\vA, \vA X+W - W'$ and outputs $W$ with probability $1-e^{-\Omega(n)}$ when $\vA$ is a random matrix.
\end{lemma}
\textbf{Note:} If we consider the setting where $t\leq m/n-2$, then $\decode_t$ will run in expected time $O(n^3\log n)$.  Our main theorems apply when $t\leq m/n-2$ with slightly better parameters.  We concentrate on the setting $t\leq d\log n(m/n-2)$.  Also even though we are correcting $O(\log n)$ errors, this a super polynomial number of error patterns and the running time of $\decode_t$ is significantly better than exhaustive search over these error patterns.

@@

\begin{theorem}
\label{thm:lossless secure extractor log}
Let $n$ be a security parameter and let $t = c\log n$ for some constant $c$.  Let $d\in \mathbb{Z}^+$ be a constant and consider the Hamming metric with block length $b = \log (2n^3 (c+2d)/d)$.  Let $W$ be uniform over $\zo^{(c+2d)nb/d}$.  If GAPSVP and SIVP are hard to approximate within polynomial factors~(on lattices of dimension $n/4$) using quantum algorithms, then there is a setting of $q = \poly(n)$ such that \consref{cons:informal construction} is a lossless fuzzy extractor.  That is, $H^{\hill}_{\epsilon, s_{sec}}(X_{1,..., n/2}|\vA, \vA X+W) = n/2\log q = H_\infty(W)$
% for following parameters:
%\[
%(\zo^{(c+d)/dn\log b}, |W|, \epsilon, s_{sec}, s_{sketch}, s_{rec}, t)
%\] 
with 
$\epsilon = \ngl(n), s_{sec} = \poly(n)$,  sketch runs in time $O(n^3\log n)$, and recover takes expected time $O(n^{4d+3} \log n)$ with error probability $\delta = e^{-\Omega(n)}$ over the choice of $\vA$.
\end{theorem}
\begin{proof}
Security follows by combining Lemmas~\ref{lem:uniform LWE decision} and~\ref{lem:many hardcore bits}, efficiency follows by \lemref{lem:i t poly time}.
\end{proof}

For a more detailed explanation of the various parameters and constraints see \secref{sec:parameter settings}.  In particular, we describe how to set $q = \poly(n)$.  
\thref{thm:lossless secure extractor log} shows that a computational fuzzy extractor can be built without incurring any entropy loss.  This is essentially due to $\vA X+W$ looking like a uniformly random point and providing no information about either $X$ or $W$.  
%A natural question to ask is whether security degrades gracefully when $W$ is not the uniform distribution.  We begin to provide an answer in \secref{sec:LWE block fixing sources}.

@@

\section{Computational Fuzzy Exactor with Nonuniform Sources}
\label{sec:LWE block fixing sources}
In this section, we show the security of~\consref{cons:informal construction} is preserved for certain high entropy distributions.  Showing security of a computational fuzzy extractor for arbitrary high min-entropy distributions is an open problem.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

We now study \consref{cons:informal construction} with a symbol fixing source.  Obviously, if $\alpha \geq n$ then the adversary knows $n$ equations and their error and $W$ can be recovered with Gaussian elimination.  We show that security degrades exponentially as symbols are fixed.
\ignore{
\begin{assumption}[LWE w/ entropic errors]
Let $n$ be a security parameter and define $m = \poly(n)$ and $q\geq 2$ be integers.  Let $W = W_1||...||W_{m+\alpha}$ be a $m$-block fixing distribution over $\{0,1\}^{(m+\alpha)\ell}$. For a vector 
Let $A\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and let $x\in\Fq^{n+\alpha}$.  No PPT algorithm given $Ax+W$ can recover $x$ with probability greater than nonnegligible.
\end{assumption}
}
We now define the decisional $\LWE$ problem with symbol fixing errors is defined as follows: 
\begin{problem}[$\distLWE$ w/ symbol fixing errors]
Let $n$ be a security parameter and define $m = \poly(n)$ and $q\geq 2$ be integers.  Let $W = W_1||...||W_{m+\alpha}$ be a $m$-symbol fixing distribution over $\{0,1\}^{(m+\alpha)b}$. For a vector $X \leftarrow \Fq^n$, we denote $A_{X, W}$ be the distribution obtained by choosing a uniform $\vA\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and outputting $\vA X + W$. The decisional learning with errors problem $\distLWE_{n+\alpha, m+\alpha, q, W}$ with symbol fixing errors is to distinguish samples chosen according $A_{X, W}$ for a uniformly random $X \in \Fq^{n+\alpha}$ from samples chosen according to the uniform distribution over $\Fq^{n+\alpha} \times \Fq$.
\end{problem}

The $\distLWE$ with symbol fixing sources is implied by standard $\distLWE$.
%\vspace{.1in}
\begin{theorem}
\label{thm:blockLWE}
Let $n$ be a security parameter and define $q, m , \alpha = \poly(n)$.  Let $\beta  = \omega(1)$, and $W$ be a $m$-symbol fixing source over $\{0,1\}^{(m+\alpha)b}$ where $\Hoo(W) \geq mb$.  The $\distLWE_{n, m,q, W}$ assumption implies the $\distLWE_{n+\alpha+\beta, m+\alpha, q, W}$ w/ symbol fixing sources problem.
\end{theorem}
We believe this statement is of interest outside our setting. \thref{thm:blockLWE} says that $\LWE$ is still secure when a small number of $\LWE$ samples have known errors.  Roughly, we need the number of variables larger than the number of equations with known error by letting the number of variables be the security parameter.  Providing a single sample with no error ``fixes'' at most a single variable.  Thus, if there are significantly more variables than samples with no error, search $\LWE$ should still be hard.  We are able to show~(a stronger result) that $\distLWE$ is still hard.  In the reduction, we argue that the additional $\alpha+ \beta$ variables allow us to explain a random value for the last $\alpha$ samples without knowledge of the other variables.  The $\beta$ parameter is the slack needed to ensure that the ``free'' variables have influence on the last $\alpha$ samples.  A formal proof is in \secref{sec:proof of block theorem}.

@@

\thref{thm:lossless block sketch log} is a direct consequence of \thref{thm:blockLWE} and parameter settings~(discussed in  \secref{ssec:block params}).  The weaker entropy source means we are able to extract fewer variables~($n/3$ instead of $n/2$) and that we must work over a slightly larger field.  We could set $\alpha$ larger than $n/3$ and extract fewer hardcore dimensions.  To extract enough bits and keep the field size polynomial in $n$, $n-\alpha$ must be at least a constant in $n$.  The parameter $\beta$ does not significantly impact the asymptotic behavior but is important for small values of $n$.

>>>>

\title{\textbf{Computational Fuzzy Extractors}}
%\author{Benjamin Fuller%\footnote{The work of Benjamin Fuller is sponsored by the United States Air Force under Air Force Contract FA8721-05-C-0002. Opinions, interpretations, conclusions and recommendations are those of the authors and are not necessarily endorsed by the United States Government.}
% \and Xianrui Meng \and Leonid Reyzin}
\author{}
\begin{document}
\maketitle

@@

\item\textbf{Positive Result:} We show that the negative result can be overcome by
analyzing computational fuzzy extractors directly.  Namely, we show
how to build a computational fuzzy extractor whose output key length
equals the entropy of the source (this is impossible in the
information-theoretic setting). Our construction is based on the
hardness of the Learning with Errors (LWE) problem, and is secure when
the noisy source is uniform or symbol-fixing (that is, each dimension
is either uniform or fixed). As part of the security proof, we show a result of independent interest, namely
that the decision version of LWE is secure even when a small number of
dimensions has no error.
\end{itemize}
\textbf{Keywords} Fuzzy extractors, secure sketches, key derivation, Learning with Errors, error-correcting codes, computational entropy, randomness extractors.
\end{abstract}

@@

The rest of this section is devoted to making these intuitive statements precise.
 We describe the \class{LWE} problem and the security of our construction in \secref{subsec:LWE}.
 %a recent variant due to D\"{o}ttling and M\"{u}ller-Quade~\cite{dottling2012} in \secref{subsec:LWE}.  
We describe one possible polynomial-time $\decode_t$ (which corrects more errors than is possible by exhaustive search) in \secref{sec:time main construction}.  In \secref{sec:lossless extractor}, we describe parameter settings that allow us to extract as many bits as the input entropy, resulting in a lossless construction.  In \secref{sec:prg based comparison}, we compare \consref{cons:informal construction} to using a sketch-and-extract approach (\lemref{lem:fuzzy ext construction}) instantiated with a computational extractor.

\subsection{Security of \consref{cons:informal construction}}
\label{subsec:LWE}
The $\LWE$ problem was introduced by Regev \cite{regev2005LWE, regevLWEsurvey} as a generalization of ``learning parity with noise." For a positive integer $n$,  a prime $q$, a vector $\vx \in \Fq^n$, and a probability distribution $\chi$ on $\Fq$, let $A_{\vect{x}, \chi}$ be the distribution obtained by choosing $\vA \overset{\$}\leftarrow \Fq^{m\times n}$ uniformly at random and a noise term $\ve \overset{\$}\leftarrow \chi^m$, and outputting $(\vA, \vA\vx+\ve)$. 
%The Learning With Errors ($\lwe$) is defined as follows: given access to the oracle that outputs a sample from the distribution $A_{\vx, \chi}$ for a uniform $\vx$, and output $\vx$ with high probability.

\begin{definition}[Decisional $\lwe$]\label{def:dist-LWE}Let $n$ be a security parameter.  
Let $m = m(n) = \poly(n)$ be an integer and $q = q(n) = \poly(n)$ be a prime. The decisional version of the $\LWE$ problem, denoted \class{dist}-$\LWE_{n, m, q, \chi}$, is to distinguish between $m$ samples chosen according to $A_{\vx, \chi}$ for a uniformly random $\vx$ from samples chosen according to the uniform distribution over $(\Fq^{m\times n}, \Fq^m)$.

We say that $\distLWE_{n, m, q, \chi}$ is $(\epsilon, s_{sec})$-secure if no (probabilistic) distinguisher of size $s_{sec}$ can distinguish the $\lwe$ instances from uniform except with probability $\epsilon$.  If for any $s_{sec} = \poly(n)$, there exists   $\epsilon  = \ngl(n)$ such that  $\distLWE_{n, m, q, \chi}$ is$(\epsilon, s_{sec})$-secure, then we say  it is \emph{secure}.
\end{definition}

 Regev\cite{regev2005LWE} and Peikert \cite{peikert2009latticereduction} show that $\class{dist}$-$\lwe_{n, m, q, \chi}$ is secure when the distribution $\chi$ of errors is Gaussian, as stated in the following proposition.
%The matrix $\vA$ and secret $\vx$ are drawn from the uniform distribution, while the error term is drawn from a discretized Gaussian distribution $\bar{\Psi}_{\alpha}$.  
%Furthermore, Regev \cite{regev2005LWE} shows a decision to search equivalence for $q=\poly(n)$.
%We state the decisional version in terms of unpredictability entropy~(see \defref{def:unp entropy} for the definition of unpredictability entropy and \secref{sec:LWEoverview} for an equivalence to the standard formulation of decisional LWE):

\begin{proposition} 
\label{assume:entropy LWE}
Let $n$ be a security parameter, let $m = m(n) = \poly(n)$ be an integer and $q = q(n) = \poly(n)$ be a prime and let $\bar{\Psi}_\rho$ be the discretized Gaussian distribution with variance $(\rho q)^2/2\pi$.  Let $\rho \in (0,1)$ with $\rho q > 2\sqrt{n}$.  If GAPSVP and SIVP are hard to approximate~(on lattices of dimension $n$) within polynomial factors for quantum algorithms, then $\distLWE_{n, m, q, \bar{\Psi}_\rho}$ is secure.

@@

\subsection{Efficiency of \consref{cons:informal construction}}
\label{sec:time main construction}
%\subsection{Analysis of Construction~\ref{cons:LWESecureSketch}}
\consref{cons:informal construction} is only useful if $\decode_t$ can be efficiently implemented.  We need a decoding algorithm for a random linear code with $t$ errors that runs in polynomial time.  We present a simple $\decode_t$.  If better decoding algorithms become known, they can replace our algorithm, which will increase our correcting capability and improve \consref{cons:informal construction}.

\newpage
\begin{construction}
\label{cons:decoding algorithm} We consider a setting of $(n, m, q, \chi)$ where $m\geq 3n$.  We describe $\decode_t$:
\begin{enumerate}
\item Input $\vA , B = \vA X + W - W'$
\item Randomly select rows without replacement $i_1,..., i_{2n}\leftarrow [1,m]$.  
\item Restrict $\vA, B$ to rows $i_1,...,i_{2n}$; denote these $\vA_{i_1,...,i_{2n}}, B_{i_1,...,i_{2n}}$.
\item Find $n$ rows of $\vA_{i_1,..., i_{2n}}$ that are linearly independent.  
If no such rows exist go to step (2).
\item Denote by $\vA', B'$ the restriction of $\vA_{i_1,..., i_{2n}}, B_{i_1,..., i_{2n}}$ (respectively) to these rows. Compute $X' = (\vA')^{-1}B'$.  
\item If $B- \vA X'$ has more than $t$ nonzero coordinates, go to step (2).
\item Output $X'$.
\end{enumerate}
\end{construction}

Each step is computable in time $O(n^3)$. 
For $\decode_t$ to be efficient, we need $t$ to be small enough so that  with probability at least $\frac{1}{\poly(n)}$, none of the $2n$ rows  selected  in step 2 have errors (i.e., so that $W$ and $W'$ agree on those rows), and $\vA_{i_1,...,i_{2n}}$ has rank  $n$.  If this happens, then $X'=X$, and the algorithm terminates.  However, we also need a second condition, to ensure correctness: we need to make sure that if $X'\neq X$, we detect it in step 6.  This detection will happen if $B-\vA X' = \vA (X-X')+(W-W')$ has more than $t$ nonzero coordinates.  It suffices to ensure that $\vA (X-X')$ has at least $2t+1$ nonzero coordinates (because at most $t$ of those can be zeroed out by $W-W'$), which happens whenever the code generated by $\vA$ has distance $2t+1$.

Setting $t = O(\frac{m}{n}\log n)$ is sufficient to ensure efficiency.    Random linear codes have distance at least $O(\frac{m}{n}\log n)$ with probability $1-e^{-\Omega(n)}$ (the exact statement is in \corref{cor:code high distance}), so this also ensures correctness.
The formal statement is below~(proof in \secref{sec:proof lem i t poly time}):
%  We provide the results in \lemref{lem:i t constant time} and \lemref{lem:i t poly time} respectively~(proofs in Sections~\ref{sec:proof lem i t constant time},~\ref{sec:proof lem i t poly time}).  \lnote{can we combine these lemmas into one?  It's hard for the reader to compare line-by-line to see where the differences are.  More errors seems better; we could simply point out that constant number of erros leads to better running time, right in the statement of the lemma or right after it}
%\begin{lemma}[Efficiency of $\decode_t$ when $t\leq m/n-1$]
%\label{lem:i t constant time}
%Let $\vA, \vA X+E$ be a  $(n,m, q, \chi)$-LWE instance that is a unique witness relation~(except with negligible probability).  Assume $d(E, E')\leq t$ where $t \leq (m-n)/n$.  Then $\decode_t$ runs in expected time $O(n^3\log q+ s_{ver})$ on input distribution $\vA, \vA X+E - E'$ and outputs $X' =X$ for all but a negligible fraction of inputs over $\vA, \vA X+E - E'$.
%\end{lemma}
\begin{lemma}[Efficiency of $\decode_t$ when $t\leq d (m/n-2)\log n$]
\label{lem:i t poly time}
%Let $\vA, \vA X+W$ be a  $\distLWE_{n,m, q, \chi}$ instance where $m\geq 3n$.  Assume that $W$ is split into $m$ blocks of length $b$ and let $\dis$ be the Hamming distance over alphabet $2^b$. 
Let $d$ be a constant and assume that $\dis(W, W')\leq t$ where $t\leq d(\frac{m}{n}-2)\log n$.  Then $\decode_t$ runs in expected time $O(n^{4d+3})$ operations in $\Fq$ (the expectation is taken over the input distribution $\vA, \vA X+W - W'$).  It outputs $X$ with probability $1-e^{-\Omega(n)}$ (this probability is over the choice of the random matrix  $\vA$).
\end{lemma}
\textbf{Note:}
Even though we are correcting $O(\log n)$ errors, this corresponds to a super polynomial number of possible error patterns; the running time of $\decode_t$ is significantly better than exhaustive search over these error patterns. Furthermore, if we consider the setting of $t\leq m/n-2$, then $\decode_t$ will run in expected time $O(n^3)$ field operations (this does not immediately follow from \lemref{lem:i t poly time}, but requires a simple reanalysis).  Thus, our main theorems get  better parameters when  $t\leq m/n-2$.

@@

\begin{theorem}
\label{thm:lossless secure extractor log}
Let $n$ be a security parameter and let $t = c\log n$ for some constant $c$.    Let $d\in \mathbb{Z}^+$ be a constant and consider the Hamming metric over the alphabet of $Z$ size $2^b$, where the  block length $b = \log (2n^3 (c+2d)/d)=O(\log n)$.  Let $W$ be uniform over $\mathcal{M}=Z^m$, where $m={(c+2d)n/d}=O(n)$.  If GAPSVP and SIVP are hard to approximate within polynomial factors (on lattices of dimension $n/4$) using quantum algorithms, then there is a setting of $q = \poly(n)$ such that for any polynomial $s_{sec}=\poly(n)$ there exists $\epsilon=\ngl(n)$ such that the following holds.
\consref{cons:informal construction} is a $(\M, bm, bm, t)$  fuzzy extractor that is $(\epsilon, s_{sec})$  hard with error $\delta = e^{-\Omega(n)}$.
Sketch takes $O(n^2)$ operations over $\Fq$, and recover takes expected time $O(n^{4d+3})$ operations over $\Fq$.
% for following parameters:
%\[
%(\zo^{(c+d)/dn\log b}, |W|, \epsilon, s_{sec}, s_{sketch}, s_{rec}, t)
%\] 
\end{theorem}
\begin{proof}
Security follows by combining Lemmas~\ref{lem:uniform LWE decision} and~\ref{lem:many hardcore bits}, efficiency follows by \lemref{lem:i t poly time}. For a more detailed explanation of the various parameters and constraints see \secref{sec:parameter settings}.  In particular, we describe how to set $q = \poly(n)$.  
\end{proof}

\thref{thm:lossless secure extractor log} shows that a computational fuzzy extractor can be built without incurring any entropy loss.  We can essentially think of $\vA X+W$ as an encryption of $X$ that where decryption works from any close $W'$.
%A natural question to ask is whether security degrades gracefully when $W$ is not the uniform distribution.  We begin to provide an answer in \secref{sec:LWE block fixing sources}.

@@

\section{Computational Fuzzy Extractor for Nonuniform Sources}
\label{sec:LWE block fixing sources}
While showing the security of~\consref{cons:informal construction} for arbitrary high-min-entropy distributions is an open problem, in this section we show it for a particular class of distributions called symbol-fixing.  First we recall the notion of a symbol fixing source~(from~\cite{KZ07}): 
\begin{definition}
Let $W = (W_1,..., W_{m+\alpha})$ be a distribution where $W_i$ takes values over $\{0,1\}^b$, we say that a distribution is a $m$-\emph{symbol fixing source} if $H_\infty(W)\geq mb$ and for all $W_i$ one of the following conditions is met:
\begin{itemize}
\item $W_i \overset{d}= U_b$
\item There exists a value $c_i$ such that $\Pr[W_i = c_i] =1$.
\end{itemize}
\end{definition}
%A block fixing source is an extension of a entropy source where each bit is independently random or fixed to larger alphabets.

\ignore{
We now study \consref{cons:informal construction} with a symbol fixing source.

\begin{assumption}[LWE w/ entropic errors]
Let $n$ be a security parameter and define $m = \poly(n)$ and $q\geq 2$ be integers.  Let $W = W_1||...||W_{m+\alpha}$ be a $m$-block fixing distribution over $\{0,1\}^{(m+\alpha)\ell}$. For a vector 
Let $A\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and let $x\in\Fq^{n+\alpha}$.  No PPT algorithm given $Ax+W$ can recover $x$ with probability greater than nonnegligible.
\end{assumption}
}
We now define the decisional $\LWE$ problem with symbol fixing errors: 
\begin{problem}[$\distLWE$ w/ symbol fixing errors]
Let $n$ be a security parameter and define $m = \poly(n)$ and $q\geq 2$ be integers.  Let $W = W_1||...||W_{m+\alpha}$ be a $m$-symbol fixing distribution over $\{0,1\}^{(m+\alpha)b}$. For a vector $X \leftarrow \Fq^n$, we denote $A_{X, W}$ be the distribution obtained by choosing a uniform $\vA\in\Fq^{(m+\alpha)\times (n+\alpha)}$ and outputting $\vA X + W$. The decisional learning with errors problem $\distLWE_{n+\alpha, m+\alpha, q, W}$ with symbol fixing errors is to distinguish samples chosen according $A_{X, W}$ for a uniformly random $X \in \Fq^{n+\alpha}$ from samples chosen according to the uniform distribution over $\Fq^{n+\alpha} \times \Fq$.
\end{problem}

The following theorem states the main technical result of this section, which is of potential interest outside our specific setting. The result is that $\distLWE$ with symbol-fixing sources is implied by standard $\distLWE$,  with security degrading exponentially as symbols are fixed (of course, once $n$ symbols are fixed, no security remains, because there are $n$ linear equations with known error and $n$ unknowns).
%\vspace{.1in}
\begin{theorem}
\label{thm:blockLWE}
Let $n$ be a security parameter and define $q, m , \alpha = \poly(n)$.  Let $\beta  = \omega(1)$, and $W$ be a $m$-symbol fixing source over $\{0,1\}^{(m+\alpha)b}$~(that is, $\Hoo(W) = mb$).  The $\distLWE_{n, m,q, W}$ assumption implies the $\distLWE_{n+\alpha+\beta, m+\alpha, q, W}$ w/ symbol fixing sources problem.
\end{theorem}

The intuition for this result is as follows.  Providing a single sample with no error ``fixes'' at most a single variable.  Thus, if there are significantly more variables than samples with no error, intuitively search $\LWE$ should still be hard.  We are able to show a stronger result that $\distLWE$ is still hard.  The nontrivial part of the reduction is using the additional $\alpha+ \beta$ variables  to ``explain'' a random value for the last $\alpha$ samples, without knowing the other variables.  The $\beta$ parameter is the slack needed to ensure that the ``free'' variables have influence on the last $\alpha$ samples.  A formal proof is in \secref{sec:proof of block theorem}.

@@

\thref{thm:lossless block sketch log} is a direct consequence of \thref{thm:blockLWE} and parameter settings~(discussed in  \secref{ssec:block params}).  The weaker entropy source means we are able to extract fewer variables~($n/3$ instead of $n/2$) and that we must work over a slightly larger field.  We could set $\alpha$ larger than $n/3$ and extract fewer hardcore dimensions.  To extract enough bits and keep the field size polynomial in $n$, $n-\alpha$ must be at least a constant fraction of $n$.  The parameter $\beta$ does not significantly impact the asymptotic behavior but is important for small values of $n$.
